# UK 2011 Census Output Area Key and Quick Statistics Import

<p>Maja Zalo&#382;nik </p>
17.6.2014

----------

This code is for the import and cleaning up of the June 2014 release from the ONS.

The dataset contains 26 Key statistics and 47 Quick Statistics, so all together that means 73 tables to be read.

Three issues are solved with this code:
- the fact that the tables have three-row-headers.
- the fact that in this release the tables have randomly from 3 to 7 empty rows appended at the end. 
- the fact that some tables also have extra empty columns appended on the right.

--------

Download the zipped folder into a tempfile.
```{r Download}
url <- "http://data.statistics.gov.uk/Census/KS_QS_OA_UK_V1.zip"
temp <- tempfile()
download.file(url, temp)
# get list of file names
file.names <- unzip(temp, list=TRUE)$Name[2:74]
table.names <- substr(file.names, 18, 22)

```


Let's check that the The 73 tables in the dataset have the same number of rows
```{r TableLenght}

table.nrows <- unlist(plyr::llply(file.names, function(n){
                length(count.fields(unzip(temp,n)))
                })) 
unique(table.nrows)

```

Interesting. They do not. I wonder why? (I asked the ONS, they don't see the problem..)

Now we check the assumption that all the tables have 3 header rows, 232,296 OA rows and between 3 to 7 extra empty rows.

```{r CheckEmptyRows}
# Check that the first four rows in the first column are all the same

first.four.rows.C1 <- lapply(file.names, function(n){
  read.csv(unzip(temp,n), nrows=4, header=FALSE, stringsAsFactors = FALSE)[1]
}
)

# if they are all the same, then the following is TRUE
length(unique(first.four.rows.C1)) == 1

```

OK, now we know there are always 3 header rows in each table. Let's make them into variable names

```{r FixHeaders}
# first uxtract only the top three rows from each table
list.all.headers <- lapply(file.names, function(n){
  read.csv(unzip(temp,n), nrows=3, header=FALSE, stringsAsFactors = FALSE)
}
)

# create a list of variable names pasted from the three header rows in each table
all.var.names<- lapply(1:73, function(i) {
  var.names <- list.all.headers[[i]][3,1]
  for (v in 2:length(list.all.headers[[i]])){
    x <- list.all.headers[[i]][v]
    nx <- stringr::str_replace_all(string=paste(x[1,], x[2,], x[3,], sep="--"), pattern=" ", repl=".")
    var.names <- c(var.names, nx)
  }
  return(var.names)
}
)

# also to keep it neat, give the list the correct names
names(all.var.names) <- table.names
```

Now we can actually import the tables and add the appropriate variable names

```{r Import tables}
all.tables <- lapply(file.names, function(n){
  read.csv(unzip(temp,n), nrows=232296, skip = 3, header=FALSE, stringsAsFactors = FALSE)
}
)
# add the variable names to each of the 73 tables
for (i in 1:73) {
  names(all.tables[[i]]) <- all.var.names[[i]]
}

# add the table names to the list
names(all.tables) <- table.names
```
head(all.tables[[1]])
all.var.names[1]
